Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2016-12-06 reresubmit/Response.txt

We would like to thank the three referees for their comments on our manuscript.  We have previously addressed the substantive concerns of Referees A and B, and in our current submission we have addressed the main concerns of Referee C. We believe that we have now addressed all of the concerns of the three referees and that our manuscript is now appropriate for publication in the Physical Review. We highlight some relevant comments from the referees in support of our manuscript:

@@

(1) It is unclear that the estimated logical error probability
reflects the true logical error probability appropriately. Suppose two
pair creations occur simultaneously nearby, two anyons are braided
with each other, and then syndrome is measured after that. If pair
annihilations are done by the decoder, the original information seems
to be restored according to the authors analysis. But it does not
return to the original one because of the braiding of two anyons. From
what reason, the authors can neglect the braiding process of anyons,
which occurs with a high probability around the threshold? How can the
authors fairly validate their simulation neglecting the braiding
dynamics of the anyons?

Response: In fact, treating the effects that the referee describes was the main challenge of our work. In the model that we used in this paper, the logical state was encoded in the non-trivial topology of the manifold. This meant that only braids that form homologically non-trivial loops will affect the encoded information. In any case, we have simulated completely all the outcomes of braids between anyons (including any effect on the logical state), which was our main technical achievement, and so we have been careful not to neglect any important effects that they may have on the error-correction process.

(2) Do the authors assume that quantum computation is done by anyonic
excitations? If so, we have to consider the dynamics between the
anyons which are employed for quantum computation and undesired
excitations. In the case of the topological codes consisting of qubits
or qudits, we can enlarge the topological defects (and hence the code
distance). However this is not the case for the "physical" anyons
encoding the logical quantum information. If a pair creation occur
nearby it and the excitation is braided around the anyon encoding the
logical quantum information, the logical information would be
destroyed. This seems to worsen the threshold.

Response: Our results apply equally well to a code where the information is stored in the topology of the manifold as to a code where the information is stored in the fusion outcomes of several non-abelian anyons, as would be the case in an anyonic quantum computer. We explicitly treated both cases in our previous work for non-universal non-abelian anyons (Phys. Rev. X 4, 031058). If the information were stored in fusion outcomes of multiple anyons, one might be concerned about braid processes where "noise" anyons braid around the "code" anyons. However, in comparison to topological defects like punctures, such a braid is not a logical error. Since the logical degree of freedom is the joint fusion channel of a pair of anyons, only braid processes that go around the entire pair will contribute to a logical error. A braid that goes around a single code anyon may interact with the charge of that anyon, but it will not interact with the logical state. This is the principal strength of topological fault tolerance for anyonic computation. In our error-correction procedure, we fuse each code anyon with any charges that lie near it, and this reverses any effects of small braids around that code anyon. The fact that one need not enlarge the size of an anyon to increase the distance of the code (as would be the case with punctures) can actually be an advantage in some circumstances by reducing the overhead.

@@

(3) While the introduction of the non-Abelian anyon system is quite
pedagogical and is well mathematically organized, the main
contribution of the manuscript, that is, the proposed efficient
algorithm to sample the non-Abelian anyons on the lattice is described
unclearly. The authors should write down the procedure more explicitly
by using equations and mathematics. The lack of the concrete
description unfortunately make it hard to justify the novelty of the
proposal.

Response: We have addressed this concern by merging the supplementary material, where the main discussion of our algorithm takes place, with the main text, together with a few additional clarifying remarks. We hope that the pseudo code description of our algorithm is a sufficient aid to help non-experts grasp the principles involved. However, we recognize that a complete reimplementation of our work would require a thorough understanding of the prior literature on quantum decoders, and that it might not be completely detailed as a standalone document. Given that other referees (nb Ref A) have commented that the paper was well written, we had to choose between adding additional pedagogical material, or trusting that non-experts would get a coarser but hopefully satisfactory understanding from the current text. Because many of the ideas in our decoder are standard in the literature, we have opted to keep our current description, and we hope this is sufficient.

In summary, we believe we have addressed all of the concerns and that the manuscript is now suitable for publication in Physical Review A.

### 2016-12-07 reresubmit/Response.txt

Response: Our results apply equally well to a code where the information is stored in the topology of the manifold as to a code where the information is stored in the fusion outcomes of several non-abelian anyons, as would be the case in an anyonic quantum computer. We explicitly treated both cases in our previous work for non-universal non-abelian anyons (Phys. Rev. X 4, 031058). If the information were stored in fusion outcomes of multiple anyons, one might be concerned about braid processes where "noise" anyons braid around the "code" anyons. However, in comparison to topological defects like punctures, such a braid is not a logical error. Since the logical degree of freedom is the joint fusion channel of a pair of anyons, only braid processes that go around the entire pair will contribute to a logical error. A braid that goes around a single code anyon may interact with the charge of that anyon, but it will not interact with the logical state. This is the principal strength of topological fault tolerance for anyonic computation. In our error-correction procedure, we fuse each code anyon with any charges that lie near it, and this reverses any effects of small braids around that code anyon. The fact that one need not enlarge the size of an anyon to increase the distance of the code (as would be the case with punctures) can actually be an advantage in some circumstances by reducing the overhead.

(3) While the introduction of the non-Abelian anyon system is quite
pedagogical and is well mathematically organized, the main
contribution of the manuscript, that is, the proposed efficient
algorithm to sample the non-Abelian anyons on the lattice is described
unclearly. The authors should write down the procedure more explicitly
by using equations and mathematics. The lack of the concrete
description unfortunately make it hard to justify the novelty of the
proposal.

Response: We have addressed this concern by merging the supplementary material, where the main discussion of our algorithm takes place, with the main text, together with a few additional clarifying remarks. We hope that the pseudo-code description of our algorithm, along with explicit description of the data structures used, is a sufficient aid to help non-experts grasp the principles involved. However, we recognize that a complete reimplementation of our work would require a thorough understanding of the prior literature on quantum decoders, and that it might not be completely detailed as a standalone document. We had to choose between adding additional pedagogical material, or trusting that non-experts would get a coarser but hopefully satisfactory understanding from the current text. Because many of the ideas in our error-correction protocol are standard in the literature, we have opted to keep our current description, and we hope this is sufficient.

### 2016-12-07 reresubmit/main text/fib2.tex
tweaks

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ Wilson loops arranged in
a square lattice tiling $\Lambda$.
These \emph{tiles} will be the observables accessible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
 $H=-\sum_{t\in \Lambda}\proj{\vac}_t\;,$ 
with $\proj{\vac}_t$ the projector to charge $\vac$ at tile $t$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space contains the two-fold degenerate codespace.

@@

In order to track the state of the system we represent the state of the fusion space with 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
Unlike previous work~\cite{Brell2013}, this creates a dynamically generated basis for the fusion space 
that allows for tensor factorization of disjoint components.
A pair-creation noise process corresponds to the 
addition of an extra curve to the lattice, supporting
two new anyons.
Following the noise processes, our simulation must 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining the curves that intersect that tile, 
and then performing $R$- and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
This procedure is discussed in more detail in Section \ref{s:technical}, and the theoretical framework underlying it is expanded in Ref.~\cite{Burton2016}.
Braiding processes that encircle 
a non-trivial loop of our (toric) manifold can also be 
treated in an analogous way, following e.g.~\cite{Pfeifer2012}.

@@

Following the results of the charge measurements, the decoder determines 
a recovery operation that involves fusion of subsets of anyons. 
These fusions can be simulated and their results calculated in the same way, 
and the output charge placed at an appropriate point in the lattice. 
This procedure is iterated until either the decoder terminates successfully, 
or the simulation itself declares failure.

@@

We consider anyon systems defined on a torus.
We endow this with a $L\times L$ square lattice of observables:
$$
    \Lambda := \bigl\{ \gamma_{ij} \bigr\}_{i,j=1,...,L}
$$
These observables are the physically accessible charge observables that will be measured during syndrome extraction, before being passed to the decoder which will in turn determine physical recovery operations to be performed on the system.

We call each $\gamma_{ij}$ a \emph{tile.}
In figures we show a small gap between the tiles but this is not meant
to reflect an actual physical gap.

@@

Syndrome extraction in this model corresponds to simulating measurements of the observables $\gamma_{ij}$, which represent the total charge contained within the corresponding tile.

In order to compute measurement outcomes for the $\gamma_{ij}$
we first need to concatenate any two curve diagrams that 
participate in the same $\gamma_{ij}.$
Because each curve has vacuum total charge this can be
done in an arbitrary way:
\begin{center}
\includegraphics[width=0.3\columnwidth ]{pic-join-pairs.pdf}
\end{center}

@@

We follow essentially the same approach as taken in \cite{Abramsky2007} 
to describe elements of a Temperley-Leib algebra, but
with some extra decorations.
The key idea is to store a \emph{word} for each tile, comprised of
the letters $\bigl<$ and $\bigr>$.
Reading in a clockwise direction around the edge of
the tile from the top-left corner,
we record our encounters with each piece-of-curve,
writing~$\bigl<$ for the first encounter, and~$\bigr>$ for the
second.
We may also encounter a dangling piece-of-curve
(the head or the tail), so we use another symbol $*$ for this.
The words for the above two tiles will then be 
$\bigl<\bigl<\bigr>\bigr>\bigl<\bigr>$ and $\bigl<\bigr>*\bigl<\bigl<\bigr>\bigr>.$
When the brackets are balanced,
each such word will correspond one-to-one with an intersection
of a curve in a tile, up to a continuous deformation of the interior of the tile.
Ie. the data structure 
will be insensitive to any continuous deformation of the interior of the tile,
but the simulation does not need to track any of these degrees of freedom.

>>>>

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ Wilson loops arranged in
a square lattice tiling $\Lambda$.
These \emph{tiles} will be the observables accessible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
 $H=-\sum_{t\in \Lambda}\proj{\vac}_t\;,$ 
with $\proj{\vac}_t$ the projector to charge $\vac$ at tile $t$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space is the two-fold degenerate codespace.

@@

In order to track the state of the system we represent the state of the fusion space with 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
Unlike previous work~\cite{Brell2013}, this creates a dynamically generated basis for the fusion space 
that allows for tensor factorization of disjoint components.
A pair-creation noise process corresponds to the 
addition of an extra curve to the lattice, supporting
two new anyons.
Following the noise processes, our simulation must 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining the curves that intersect that tile, 
and then performing $R$- and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
This procedure is discussed in more detail in Section \ref{s:technical}, and the theoretical framework underlying it is expanded in Ref.~\cite{Burton2016}.
Braiding processes that encircle 
a non-trivial loop of our (toric) manifold can also be 
treated in an analogous way, following e.g.~Ref.~\cite{Pfeifer2012}.

@@

Following the results of the charge measurements, the decoder determines 
a recovery operation that involves fusion of subsets of anyons. 
These fusions can be simulated and their results calculated in the same way, 
and the output charge placed at an appropriate point in the lattice. 
This procedure is iterated until either the decoder terminates successfully, 
or the simulation itself declares failure because the components become too large to analyze (roughly when they span the lattice).

@@

We consider anyon systems defined on a torus.
We endow this with a $L\times L$ square lattice of observables:
$$
    \Lambda := \bigl\{ \gamma_{ij} \bigr\}_{i,j=1,...,L}
$$
These observables are the physically accessible charge observables that will be measured during syndrome extraction, before being passed to the decoder which will in turn determine physical recovery operations to be performed on the system.
We call each $\gamma_{ij}$ a \emph{tile.}
In figures we show a small gap between the tiles but this is not meant
to reflect an actual physical gap.

@@

Syndrome extraction in this model corresponds to simulating measurements of the observables $\gamma_{ij}$, which represent the total charge contained within the corresponding tile.
In order to compute measurement outcomes for the $\gamma_{ij}$
we first need to concatenate any two curve diagrams that 
participate in the same $\gamma_{ij}.$
Because each curve has vacuum total charge this can be
done in an arbitrary way:
\begin{center}
\includegraphics[width=0.3\columnwidth ]{pic-join-pairs.pdf}
\end{center}

@@

We follow essentially the same approach as taken in \cite{Abramsky2007} 
to describe elements of a Temperley-Leib algebra, but
with some extra decorations.
The key idea is to store a \emph{word} for each tile, comprised of
the letters $\bigl<$ and $\bigr>$.
Reading in a clockwise direction around the edge of
the tile from the top-left corner,
we record our encounters with each piece-of-curve,
writing~$\bigl<$ for the first encounter, and~$\bigr>$ for the
second.
We may also encounter a dangling piece-of-curve
(the head or the tail), so we use another symbol $*$ for this.
The words for the above two tiles will then be 
$\bigl<\bigl<\bigr>\bigr>\bigl<\bigr>$ and $\bigl<\bigr>*\bigl<\bigl<\bigr>\bigr>.$
When the brackets are balanced,
each such word will correspond one-to-one with an intersection
of a curve in a tile, up to a continuous deformation of the interior of the tile,
i.e.~the data structure 
will be insensitive to any continuous deformation of the interior of the tile,
but the simulation does not need to track any of these degrees of freedom.
